---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
# Reproducible Research: Peer Assessment 1
## Loading and preprocessing the data. 
### First unzip activity.zip file and do some basic data exploration.
```{r,echo=TRUE}
unzip("activity.zip")
library("dplyr")
data=read.csv2("activity.csv",sep=",",header=TRUE)
head(data)
names(data)
dim(data)
```

## What is mean and median total number of steps taken per day?
#### For this part of the assignment, you can ignore the missing values in the dataset. Make a histogram of the total number of steps taken each day.Calculate and report the mean and median total number of steps taken per day

```{r,echo=TRUE,results='asis'}

daily_step_totals=sapply(split(data$steps,data$date),sum)
hist(daily_step_totals,breaks=c(seq(0,25000,2500)),xlab="Total number of daily steps",ylab="Number of days",main="Histogram of total number of daily steps",col="lightgreen")
mean_daily_steps=mean(daily_step_totals,na.rm=TRUE)
median_daily_steps=median(daily_step_totals,na.rm=TRUE)
cat(paste("Mean Daily Steps:", mean_daily_steps ))
cat(paste("Median Daily Steps:", median_daily_steps ))    
```


## What is the average daily activity pattern?

```{r,echo=TRUE}
avg_daily_pattern=sapply(split(data$steps,data$interval),mean,na.rm=TRUE)
plot(ts(avg_daily_pattern),type="l",col="red",ylab="Average number of steps",xlab="Time of Day")
hours=data$interval %/% 100
minutes=data$interval %% 60

avg_daily_pattern[avg_daily_pattern==max(avg_daily_pattern)]
```
#### As would be expected, activity level drops after 10 pm and does not pick up till 6 am. Peak activity is between 8:30-8:35 am.

## Imputing missing values.

```{r,echo=TRUE,results='asis'}
tot_num_of_missing_values=sum(is.na(data))
cat(paste("Total number of missing values:",tot_num_of_missing_values  ))    
```
#### Simplest case: Impute missing values with average for that interval.Do this using if else and select and regenerating a histogram for the new dataset.

```{r,echo=TRUE,results='asis'}
means_vec=means_vec=as.vector(avg_daily_pattern)
int_vec= as.numeric(levels(as.factor(data$interval)) )
data$updated_steps=ifelse((is.na(data$steps)),means_vec[match(data$interval,int_vec)],data$steps)
updated_data=select(data,-steps)
aa=group_by(updated_data,date)
updated_stepsSum=summarize(aa,daily_tot=sum(updated_steps))
hist(updated_stepsSum$daily_tot,breaks=c(seq(0,25000,2500)),xlab="Total number of daily steps",ylab="Number of days",main="Histogram of total number of daily steps",col="lightblue")
imputed_mean_daily_steps=mean(updated_stepsSum$daily_tot,na.rm=TRUE)
imputed_median_daily_steps=median(updated_stepsSum$daily_tot,na.rm=TRUE)
cat(paste("Mean Daily Steps:", imputed_mean_daily_steps ))
cat(paste("Median Daily Steps:", imputed_median_daily_steps ))    
```
#### There is no differnce in the mean number of steps (which makes sense as I imputed the missing steps values with the average number of steps for that interval). There is a slight uptick in the median number of steps and now the median number of steps =mean number of daily steps.  

## Are there differences in activity patterns between weekdays and weekends?
```{r,echo=TRUE,results='asis'}



```
